{\enAbstractStyle
\noindent\textbf{ABSTRACT:} As the installed capacity of renewable energy sources such as wind power and photovoltaics continues to grow, power systems are shifting from a structure dominated by stable generation to one characterized by high variability, multiple participants, and strong uncertainty. Forecasting errors in renewable generation directly affect reserve allocation, energy storage scheduling, electricity trading plans, and the economic operation of microgrids. As a result, forecasting has become a fundamental capability in smart energy systems. Owing to its strong ability to model nonlinear relationships and extract temporal patterns, deep learning has been widely applied in load forecasting, wind power forecasting, photovoltaic power forecasting, and microgrid energy management in recent years. This report examines how deep learning supports smart energy applications, with a primary focus on photovoltaic power forecasting, and further discusses wind power forecasting and microgrid dispatch to summarize the main application pathways, representative models, and engineering value of deep learning in the energy domain. Overall, models such as LSTM, GRU, and CNN-LSTM show strong potential for renewable energy forecasting. However, the practical deployment of deep learning in smart energy systems still faces challenges related to data quality, cross-scenario generalization, interpretability, and deployment cost. Future research should place greater emphasis on integrating physical mechanisms with data-driven modeling, coupling forecasting with optimization, and multimodal collaborative modeling. Finally, an experiment was conducted on the public LIGHT photovoltaic dataset. In this experiment, 1-minute power data were resampled to 45-minute intervals, temporal features were added, and complete day-night sequences were preserved to compare Persistence, MLP, and LSTM in a 90-minute-ahead forecasting task. The results show that under this setting, LSTM outperforms Persistence and MLP on MAE, RMSE, and MAPE. Its RMSE reaches 0.3630, representing an improvement of about 13.28\% over Persistence. This suggests that in real photovoltaic scenarios, when the natural advantage of a simple persistence baseline is weakened by a coarser time granularity and a longer forecasting horizon, LSTM can better capture temporal dependencies. The experimental code is available at \url{https://github.com/glance02/SmartEnergyPVForecasting}.\par}
\vspace{6pt}

{\enAbstractStyle
\noindent\textbf{KEY WORDS:} smart energy systems; deep learning; photovoltaic power forecasting; wind power forecasting; microgrid energy management\par}
\vspace{6pt}

{\zhAbstractStyle
\noindent{\heiti 摘要：} 随着风电、光伏等新能源装机容量持续增长，电力系统正在从以稳定电源为主的结构转向高波动、多主体、强不确定性的运行形态。新能源出力预测误差会直接影响备用容量配置、储能调度、购售电计划以及微电网经济运行，因此预测能力已经成为智慧能源系统中的关键基础能力。深度学习由于具备较强的非线性建模能力和时序模式提取能力，近年来被广泛用于负荷预测、风电功率预测、光伏功率预测和微电网能量管理等场景。在这份报告中，我围绕“深度学习如何支撑智慧能源”这一问题展开分析，重点聚焦光伏功率预测，并结合风电预测与微电网调度案例，梳理深度学习在智慧能源中的主要应用路径、代表性模型和工程价值。综合来看，LSTM、GRU、CNN-LSTM 等模型在新能源预测中具有较好应用前景，但深度学习在智慧能源中的落地仍面临数据质量、跨场景泛化、可解释性和部署成本等问题。未来研究应更多关注机理与数据融合、预测与优化一体化以及多模态数据协同建模等方向。最后，我基于公开的 LIGHT 光伏数据集做了一个实验：将 1 分钟功率数据重采样为 45 分钟数据，加入时间周期特征，并保留完整昼夜序列，比较 Persistence、MLP 和 LSTM 在 90 分钟 ahead 预测任务中的表现。实验结果表明，在这一设定下 LSTM 在 MAE、RMSE 和 MAPE 三项指标上都优于 Persistence 和 MLP，其中 RMSE 为 0.3630，较 Persistence 下降约 13.28\%。这说明在真实光伏场景中，若通过更粗时间粒度和更长预测步长削弱简单延续性基线的天然优势，LSTM 对时序依赖关系的建模能力能够更好体现出来。我的小实验代码开源在 \url{https://github.com/glance02/SmartEnergyPVForecasting}。\par}
\vspace{6pt}

{\zhAbstractStyle
\noindent{\heiti 关键词：} 智慧能源；深度学习；光伏功率预测；风电预测；微电网能量管理\par}
\vspace{6pt}
